<!-- README.md is generated from README.rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# chents

[![Online documentation](https://img.shields.io/badge/docs-jrwb.de-blue.svg)](https://pkgdown.jrwb.de/chents/)
[![R-Universe status](https://jranke.r-universe.dev/badges/chents)](https://jranke.r-universe.dev/chents)
[![Code coverage](https://img.shields.io/badge/coverage-jrwb.de-blue.svg)](https://pkgdown.jrwb.de/chents/coverage/coverage.html)

The R package **chents** provides some utilities for working with chemical
entities in R.

When first defining a chemical entity, some chemical information
is retrieved from the [PubChem](https://pubchem.ncbi.nlm.nih.gov/) website using
the [webchem](https://docs.ropensci.org/webchem/) package.

```{r}
library(chents)
caffeine <- chent$new("caffeine")
```

If Python and [RDKit](https://rdkit.org) (> 2015.03) are installed and
configured for use with  the
[reticulate](https://rstudio.github.io/reticulate/) package, some
additional chemical information including a 2D graph are computed.

The print method gives an overview of the information that was collected.

```{r}
print(caffeine)
```

There is a very simple plotting method for the chemical structure.

```{r fig.width = 2}
plot(caffeine)
```

Additional information can be (but is rarely ever) read from a local .yaml
file. This information can be leveraged e.g. by the
[PEC_soil](https://pkgdown.jrwb.de/pfm/reference/PEC_soil.html) function of the
'pfm' package.

If you have a so-called ISO common name of a pesticide active ingredient, you
can use the 'pai' class derived from the 'chent' class, which starts with querying
the [BCPC compendium](http://www.bcpcpesticidecompendium.org/) first.

```{r fig.width = 2}
lambda <- pai$new("lambda-cyhalothrin")
plot(lambda)
```


## Installation

You can conveniently install chents from the repository kindly made available by the
R-Universe project:

```
install.packages("chents",
  repos = c("https://jranke.r-universe.dev", "https://cran.r-project.org"))
```

In order to profit from the chemoinformatics, you need to install RDKit and its
python bindings. On a Debian type Linux distribution, just use

```
sudo apt install python3-rdkit
```
If you use this package on Windows or MacOS, I would be happy to include
installation instructions here if you share them with me, e.g. via a Pull
Request.

## Configuration of the Python version to use

On Debian type Linux distributions, you can use the following line in your
global or project specific `.Rprofile` file to tell the `reticulate` package to
use the system Python version that will find the RDKit installed in the system
location.

```
Sys.setenv(RETICULATE_PYTHON="/usr/bin/python3")
```

